A geo-referenced micro-data set of real estate listings for Spain's three largest cities

### 2023-05-19 editorial-materials/EPB-2022-0116.R1_response_to_reviewer.Rmd

---
title: "A geo-referenced micro-data set of real estate listings for Spain’s three largest cities - Response to referees"
output: pdf_document
bibliography: bibEPB.bib
link-citations: yes
linkcolor: blue
---

```{r load-packages, include=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
library(gridExtra)
library(idealista18)
library(kableExtra)
library(sf)
```

@@

>4. For data description, it is worth showing solid evidence of the quality of this dataset. For example, what is the data coverage of this data compared to all the residential properties in the city?  Listing the advertised volume is not very helpful for users who wish to understand the quality.

**Response:**

To better place in context the coverage of the {idealista18} dataset, Table \ref{tab:transactions} now shows the number of listings with respect to the total residential stock in each city (in 2018). As seen in the table, the number of listings ranges between 6.1% of the total number of properties (in Madrid) and 8.1% (in Valencia). Information from Instituto Nacional de Estadistica^[\url{https://www.ine.es/jaxiT3/Tabla.htm?t=6150&L=1}] for 2018 also shows that the number of transactions in 2018 are equivalent to 81.3\% of idealista listings in Barcelona, 80.8\% in Madrid, and 91.1\% in Valencia. While it is not possible to link actual transactions to idealista listings, this gives a quantitative sense of the coverage of the data package.

```{=tex}
\begin{table}[!ht]
\centering
\caption{Total properties and transactionst three Spanish cities. Year 2018} 
\label{tab:transactions} 
\begin{tabular}{lccccc}
City & Total properties (TP) & Total transactions (TT) & Listings (L) & L/TP & TT/L \\
\hline
Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
\hline
Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
\hline
\multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
\multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
\multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
\end{tabular}
\end{table}
```

@@

**Response:**

We appreciate the opportunity to delve more deeply into this issue, as it may be of interest to users of the data package. The short version of this story is that adding a small amount of random noise to the prices ($\pm 2.5\%$) does not introduce bias but does increase the variance of the masked variable. This can be shown with some standard algebra of random variables.

Now the longer story.

The masked prices are given by:
$$
P = RP\cdot \epsilon
$$

\noindent where $\epsilon$ is drawn from the uniform distribution with parameters $a=0.975$ and $b=1.025$:
$$
f(\epsilon) = 
\begin{cases}
\frac{1}{b - a} & \text{for } a\le \epsilon \le b\\
0 & \text{otherwise}
\end{cases}
$$

The expectation of $\epsilon$ given these parameters is:
$$
E[\epsilon] = \frac{a+b}{2} = \frac{0.975 + 1.025}{2} = 1
$$

\noindent and the variance of $\epsilon$ is:
$$
\mathrm{V}[\epsilon]=\frac{(b - a)^2}{12} = \frac{1}{4800}
$$

Therefore, the expectation of the masked prices is:
$$
\mathrm{E}[P] = \mathrm{E}[RP\cdot \epsilon] = \mathrm{E}[RP]\cdot \mathrm{E}[\epsilon] = \mathrm{E}[RP]
$$

In other words, the masked prices $P$ are an unbiased version of the raw prices $RP$.

Considering that $RP$ and $\epsilon$ are independent, the variance of the masked prices is as follows:
$$
\mathrm{V}[P] = \mathrm{V}[RP\cdot \epsilon] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP]\cdot (\mathrm{E}[\epsilon])^2 + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
$$

Since $\mathrm{E}[\epsilon] = 1$, we have that:
$$
\mathrm{V}[P] = \mathrm{V}[RP]\cdot \mathrm{V}[\epsilon] + \mathrm{V}[RP] + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2 = \mathrm{V}[RP]\cdot (1 + \mathrm{V}[\epsilon]) + \mathrm{V}[\epsilon]\cdot(\mathrm{E}[RP])^2
$$

Solving for $\mathrm{V}[RP]$, and replacing the expectation of the raw prices by its unbiased version ($\mathrm{E}[P]$), yields the variance of the raw prices:
$$
\mathrm{V}[RP] = \frac{\mathrm{V}[P] - \frac{1}{4800}\cdot (\mathrm{E}[P])^2}{1 + \frac{1}{4800}}
$$

Table \ref{tab:variance} reports the mean and the standard deviation of the prices in the package, and the standard deviation of the raw prices calculated using the formula above. The last column of the table can be read as an inflation factor. We see that the variance of the masked prices is inflated with respect to the variance of the original values by less than $1\%$ in all cases examined. Users can use the formula above to calculate the inflation of the variance if they use subsamples other than those shown, to assess the potential impacts of the masking (e.g., when computing intervals of confidence).

```{r echo = FALSE}
data("Barcelona_Sale")
data("Madrid_Sale")
data("Valencia_Sale")
```

```{r echo=FALSE}
# Barcelona
Barcelona_Price <- Barcelona_Sale |>
  select(PRICE, PERIOD) |>
  st_drop_geometry()

BQ <- Barcelona_Price |>
  group_by(PERIOD) |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "BARCELONA")

BT <- Barcelona_Price |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(PERIOD = "2018",
         sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "BARCELONA")

# Madrid
Madrid_Price <- Madrid_Sale |>
  select(PRICE, PERIOD) |>
  st_drop_geometry()

MQ <- Madrid_Price |>
  group_by(PERIOD) |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "MADRID")

MT <- Madrid_Price |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(PERIOD = "2018",
         sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "MADRID")

# Valencia
Valencia_Price <- Valencia_Sale |>
  select(PRICE, PERIOD) |>
  st_drop_geometry()

VQ <- Valencia_Price |>
  group_by(PERIOD) |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "VALENCIA")

VT <- Valencia_Price |>
  summarize(mean = mean(PRICE),
            variance = var(PRICE),
            .groups = "drop") |>
  mutate(PERIOD = "2018",
         sq_mean = mean^2,
         var_raw = (variance - sq_mean/4800)/(1 + 1/4800),
         inflation = variance/var_raw,
         CITY = "VALENCIA")

Q <- rbind(BQ,
           BT,
           MQ,
           MT,
           VQ,
           VT)

Q |>
  mutate(PERIOD = case_when(PERIOD == "201803" ~ "Q1",
                            PERIOD == "201806" ~ "Q2",
                            PERIOD == "201809" ~ "Q3",
                            PERIOD == "201812" ~ "Q4",
                            PERIOD == "2018" ~ "2018"),
         sd = round(sqrt(variance), 2),
         sd_raw = round(sqrt(var_raw), 2),
         inflation = round(inflation, 5),
         across(c(mean, sd, sd_raw, inflation), ~ prettyNum(.x, big.mark = ",")))  |>
  select(CITY, PERIOD, mean, sd, sd_raw, inflation) |>
  kableExtra::kable(booktabs = TRUE,
                    caption = "\\label{tab:variance}Inflation of the variance of masked prices with respect to raw prices",
                    col.names = c("City", "Period", "mean(P)", "sd(P)", "sd(RP)", "sd(P)/sd(RP)")) |>
  collapse_rows(columns = 1) |>
  footnote(general = c("P: masked prices; ", "RP: raw prices; ", "sd: standard deviation (square root of the variance)"),
           threeparttable = T)
```

@@

>6. In figure 1, is it possible to standardise this data rather than showing only the number of dwelling records in your house price data? I guess the city centre always has high advertised volumes but it also has a high density of dwellings. It is worthwhile to show values standardised by the number of dwellings.

**Response:**

After this comment we included a new figure in the paper to display the percentage of listings relative to the total number of dwellings by district (see Figure \ref{fig:districts} below).

```{r, echo = FALSE,message=FALSE, fig.height=8, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
data("properties_by_district")

### 2023-05-19 EPB/main_EPB.Rmd
Resubmission to EPB

---
title: A geo-referenced micro-data set of real estate listings for Spain's three largest cities
runninghead: author \emph{et al}.
author:
- name: 
abstract: "This article presents an open data product with large geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented R package and is available for scientific and educational purposes, particularly for geo-spatial studies"

@@

Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature. Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market: to this end, it is essential to have information/data at the point level. Therefore, it is becoming common for spatial analysis of urban environments to be use micro-data, geo-referenced as points [@lopez2015]. In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @gupta2022take; @arbia2020spatial; @Li2019; @lopez2015]. Alas, webscraping is a chancy process prone to download errors, missing  data, duplicated records, etc. Furthermore, researchers do not always share their webscrapped datasets, which limits reproducibility of their research. As we witness a growing interest in openness and reproducibility in geographical data science [@arribasl2021editorial; @paez2021open; @brunsdon2021opening], it  becomes increasingly urgent to have open data products to support research [@arribas2021].

@@

To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open micro-data set of geo-referenced dwelling listings. The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain's three largest cities. To date, this data product is one of the largest open geo-referenced micro-data set of the housing market in the world. The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$). It is certainly the largest in Spain.

The data set has been supplied directly by Idealista, and therefore is clean and free of download errors. However, to comply with data legislation, to share the data we have slightly modified the prices by applying a small amount of random noise which will not bias the main results derived from its usage.

This micro-data set can be used to benchmark new analytical methods in a reproducible fashion. Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others. The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices. The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest. In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.

The data set is distributed as an `R` package, named {idealista18}, which can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}. The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the randomization process applied to the original data in the aforementioned Github repository.

@@

```{=tex}
\begin{table}[!ht]
    \centering
      \caption{Total properties and transactionst three Spanish cities. Year 2018} 
  \label{tab:transactions} 
    \begin{tabular}{lccccc}
        City & Total properties (P) & Total transactions (T) & Listings (L) & L/P & T/L \\
        \hline
        Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
        Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
        Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
        \hline
        Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
        \hline
        \multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
        \multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
        \multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
    \end{tabular}
\end{table}
```

The following subsections describe each object. A full description of the data is available in the help section of the package. We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the R ecosystem mean that documentation is easily searchable; distribution as an R package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.

@@

The dwelling listing of each city includes a set of characteristics for each dwelling published on the idealista real estate website as an ad. The dwelling listing has been included in the {idealista18} package as a simple features (sf) object [@Pebesma] with point geometry in latitude and longitude. The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., Madrid_Sale). Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018. Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter. The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.

@@

Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including the location of the dwelling. The exact location has been masked, as described in Section [Anonymizing the data set](#anonymizing)). Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable. The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name. The cadastral features were assigned by applying the features of the nearest parcel to the coordinates. The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data). To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.

@@

```{r table-variables, echo = FALSE, message = FALSE}
dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("Asking price","Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Construction year (advertiser)","Construction year (cadastre)",
                       "Max build floor","Dwelling count in the building",
                       "Cadastral quality. 0 Best-10 Worst",
                       "Distance to city center","Distance to subway station","Distance to major street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
check.names = FALSE)
text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities. See the help section in the \\textbf{idealista18} R package for details and formal definitions. Some variables have been excluded from this table to save space. Check the full list in the \\textbf{idealista18} package.\\label{tab:variables}") |>
xtable2kable(include.rownames = FALSE) |>
kable_styling(full_width = F, font_size=8) |>
column_spec(1, width = "13em", italic = FALSE) |>
column_spec(2, width = "14em") |>
column_spec(3, width = "2em", latex_column_spec = "c") |>
column_spec(4, width = "2em", latex_column_spec = "c") |>
column_spec(5, width = "2em", latex_column_spec = "c")
```

@@

The second data object included in {idealista18} is the neighborhoods in the three cities as polygons. There is an sf object for each city with the name of the city and the suffix '_Polygons'. Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities. The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division. If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}. In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area. In the case of Madrid, they combined four areas into two.

```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:all-polygons}Number of listings in each neighborhood. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}

@@

barcelona_listings <- ggplot(data = Barcelona_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

madrid_listings <- ggplot(data = Madrid_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

valencia_listings <- ggplot(data = Valencia_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1,0.1,0.1,0.1), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

# Arrange plots
grid.arrange(barcelona_listings, madrid_listings, valencia_listings,
             ncol = 1,
             nrow = 3)
```

@@

The total number of dwellings is available from the Spanish cadastre aggregated by district (districts are groups of neighborhoods). Figure \ref{fig:districts} shows the percentage of listed dwellings relative to the total number of dwelling by district in each of the three cities. This gives a sense of how active residential real estate markets were in different parts of each city in 2018.

```{r, echo = FALSE,message=FALSE, fig.width=5, fig.cap="\\label{fig:districts}Percentage of listings relative to total number of dwellings by district. Boundary for Barcelona (Top), Madrid (Center), and Valencia (Bottom)."}
# Barcelona
Barcelona_prop <- properties_by_district |>
  filter(CITY == "Barcelona") |>
  mutate(proportion = N/N_CADASTRE)

@@

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA) +
  geom_sf(data = Barcelona_Polygons,
          fill = NA) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.1, 0.1, 0.1),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

# Anonymizing the data set {#anonymizing}

To comply with Spanish regulations, two variables were slightly modified to provide anonymity. A masking process was applied to asking prices and location (coordinates).

@@

```{r, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
```

>>>>

---
title: A geo-referenced micro-data set of real estate listings for Spain's three largest cities
runninghead: author \emph{et al}.
author:
- name: 
abstract: "This article presents an open data product with large geo-referenced micro-data sets of 2018 real estate listings in Spain. These data were originally published on the idealista.com real estate website. The observations were obtained for the three largest cities in Spain: Madrid (n = 94,815 observations), Barcelona (n = 61,486 observations), and Valencia (n = 33,622 observations). The data sets include the coordinates of properties (latitude and longitude), asking prices of each listed dwelling, and several variables of indoor characteristics. The listings were enriched with official information from the Spanish cadastre (e.g., building material quality) plus other relevant geographical features, such as distance to urban points of interest. Along with the real estate listings, the data product also includes neighborhood boundaries for each city. The data product is offered as a fully documented `R` package and is available for scientific and educational purposes, particularly for geo-spatial studies"

link-citations: yes
linkcolor: blue

@@

Interest in the characteristics of the housing market and housing prices has been a growing area of research in recent decades, generating a vast amount of theoretical and empirical literature. Including the spatial component to analyze the real estate market and incorporating geographic variables has significantly improved the understanding of this market: to this end, it is essential to have information/data at the point level. Therefore, it is becoming common for spatial analysis of urban environments to be use micro-data, geo-referenced as points [@lopez2015]. In some cases, researchers have had to resort to webscraping processes to obtain data for research [e.g., @Li2019; @lopez2015]. Alas, webscraping is a chancy process prone to download errors, missing  data, duplicated records, etc. Furthermore, researchers do not always share their webscrapped datasets, which limits reproducibility of their research. As we witness a growing interest in openness and reproducibility in geographical data science [@arribasl2021editorial; @paez2021open; @brunsdon2021opening], it  becomes increasingly urgent to have open data products to support research [@arribas2021].

@@

To contribute to the growing inventory of international micro-data sets of real estate markets, this paper introduces an open data product with geo-referenced dwelling listings, called {idealista18}. The data have been provided by the Idealista company\footnote{Idealista is the major real estate listing website in Spain, and present in other southern european countries as Italy and Portugal} and contain information about 189,923 dwellings located in Spain's three largest cities. To date, this data product is one of the largest open geo-referenced micro-data set of the housing market in the world. The most similar in terms of geographical disaggregation and sample size is the dataset of @song2021hedonic which includes transactions for four cities in South Korea, namely Busan ($n=61,152$), Daegu ($n=32,363$), Daejeon ($n=21,114$), and Gwangju ($n=25,984$). {idealista18} is certainly the largest open data product of its kind in Spain.

The data set has been supplied directly by Idealista, and therefore is clean and free from download errors. However, to comply with data legislation, we have masked the prices by applying a small amount of random noise that will not bias the main results derived from its usage.

This micro-data set can be used to benchmark new methods in a reproducible fashion  [e.g., @rey2023using]. Applied and theoretical researchers on real estate mass appraisal and valuation methods might use this dataset to canonically compare the performance of their proposed comparable and hedonic models, among others. The data can also be used to study the segmentation of housing submarkets and related topics such as the impact of suburban areas on house prices. The listings have been enriched with official information from the Spanish cadastre along with other relevant geographical features, such as distance to urban points of interest. In any case, the data might be easily extended by spatially joining other datasets that contain information at administrative, census tract, or postal code levels.

The data set is distributed as an `R` package and it can be accessed from a public Github repository\footnote{\url{https://github.com/paezha/idealista18}}. The open data product is made available under the \href{http://opendatacommons.org/licenses/odbl/1.0/}{Open Database License}. For transparency, we also share the masking process applied to the original data in the aforementioned Github repository.

@@

```{=tex}
\begin{table}[!ht]
\centering
\caption{Total properties and transactionst three Spanish cities. Year 2018} 
\label{tab:transactions} 
\begin{tabular}{lccccc}
City & Total properties (TP) & Total transactions (TT) & Listings (L) & L/TP & TT/L \\
\hline
Barcelona &  789,740  &  56,012  &  61,329  & 7.8\% & 81.3\%  \\ 
Madrid &  1,545,397  &  76,603  &  94,802  & 6.1\%  & 80.8\% \\ 
Valencia &  416,004  &  30,615  &  33,593  & 8.1\% & 91.1\% \\ 
\hline
Total &  2,751,141  &  163,230  &  189,724  & 6.9\% & 86.0\% \\ 
\hline
\multicolumn{6}{l}{{\footnotesize \textbf{Sources}}}\\
\multicolumn{6}{l}{{\footnotesize Total properties (P): Ministerio Español de Hacienda y Función Pública}}\\
\multicolumn{6}{l}{{\footnotesize Total transactions (T): Instituto Nacional de Estadística}}\\
\end{tabular}
\end{table}
```

The following subsections describe the data objects. A full description of the data is also available in the help section of the package. We have tried to the best of our ability to comply with FAIR principles regarding research data [@Wilkinson2016fair]: upon publication, the dataset has a persistent digital object identifier; publication as a data article makes the data findable; the data and metadata are packaged together and protocols for help files in the `R` ecosystem mean that documentation is easily searchable; distribution as an `R` package means that only open software is needed to access the data; and a public repository documents all data processes followed to generate the distributed open data product.

@@

The listing for each city includes a set of characteristics for each dwelling published on the idealista real estate website. The listing included in the {idealista18} package are simple features (sf) objects [@Pebesma] with point geometry in latitude and longitude. The name of each sf object with the list of dwellings is the name of the city followed by '_Sale' (e.g., `Madrid_Sale`). Each data object includes a total of 42 variables and the complete set of listings corresponding to the four quarters of 2018 (Q1 through Q4). Table \ref{tab:number-ads} shows the number of dwelling listing ads included in the data set for each city and quarter. The record counts for each city are: 94,815 listings for Madrid, 61,486 for Barcelona, and 33,622 for Valencia. Note that the same dwelling may be found in more than one period when a property listed for sale in one quarter was sold in a subsequent quarter. The variable ASSETID, included in the sf objects, is the unique identifier of the dwelling.

@@

Each record of the dwelling listing contains a set of indoor characteristics supplied by the advertisers on the Idealista website (e.g., price, surface area, number of bedrooms, basic features, etc.), including an approximated location of the dwelling (the exact location has been masked, as described in Section [Masking the prices](#anonymizing)). Table \ref{tab:variables} lists the main indoor variables with a short description and the mean value of each variable. The dwelling listings were enriched with a number of additional attributes from the Spanish cadastre [@Catastro]. The cadastral information is described in Table \ref{tab:variables}, with the prefix CAD in the variable name. The cadastral features were assigned by applying the features of the nearest parcel to the coordinates. The year the dwelling was built (CONSTRUCTIONYEAR) given by the advertiser was revised since the year of construction is entered on the website by users, and it is therefore subject to errors and incomplete data (40\% missing data). To resolve this issue, an alternative variable (CADCONSTRUCTIONYEAR) was included, assigning the cadastral construction year from the nearest cadastral parcel whenever the value was outstanding (date was after publication date or year of construction was before 1500) or when the value supplied by the advertiser was missing.

@@

```{r table-variables, echo = FALSE, message = FALSE}
dummyvariables <- c(3,4,5,6,7,27,30,31,32,33,37,38,39)
text_tbl <- data.frame(
Variable = names(st_drop_geometry(Barcelona_Sale[dummyvariables])),
"Sort Description" = c("Asking price","Asking price per m^2 (euros)","Surface (m^2)",
                       "Number of bedrooms","Number of bathrooms",
                       "Construction year (advertiser)","Construction year (cadastre)",
                       "Max build floor","Dwelling count in the building",
                       "Cadastral quality. 0 Best-10 Worst",
                       "Distance to city center","Distance to subway station","Distance to major street"),
# Barcelona
Barcelona = colMeans(st_drop_geometry(Barcelona_Sale[dummyvariables]), na.rm = TRUE),
# Madrid
Madrid = colMeans(st_drop_geometry(Madrid_Sale[dummyvariables]), na.rm = TRUE),
# Valencia
"Valencia"=colMeans(st_drop_geometry(Valencia_Sale[dummyvariables]), na.rm = TRUE),
check.names = FALSE)
text_tbl[dim(text_tbl)[1],1]<-"DISTANCE_TO_(MAINSTREET)"
xtable::xtable(text_tbl, caption = "List, sort description, and mean of the main quantitative variables included in the dwelling listing for the three Spanish cities. See the help section in the \\{idealista18\\} package for details and formal definitions. Some variables have been excluded from this table to save space. Check the full list in the package.\\label{tab:variables}") |>
xtable2kable(include.rownames = FALSE) |>
kable_styling(full_width = F, font_size=8) |>
column_spec(1, width = "13em", italic = FALSE) |>
column_spec(2, width = "14em") |>
column_spec(3, width = "2em", latex_column_spec = "c") |>
column_spec(4, width = "2em", latex_column_spec = "c") |>
column_spec(5, width = "2em", latex_column_spec = "c")
```

@@

The second data object included in {idealista18} is the neighborhoods in the three cities as polygons. There is an sf object for each city with the name of the city and the suffix '_Polygons'. The left column of Figure \ref{fig:all-polygons} shows the quantile maps of the number of dwellings in the listing for the different neighborhoods in the three cities. The neighborhoods are based on the official boundaries but slightly changed by Idealista\footnote{There are two criteria used to make this division. If an area is small enough and similar enough to another, the two areas are merged, and, if the official area is not homogeneous, it is divided into a series of new polygons.}. In practical terms, we can assume they are the same since the website combines areas when there are few ads for that area. In the case of Madrid, they combined four areas into two.

```{r listings-neighborhood, include = FALSE}

@@

plot1 <- ggplot(data = Barcelona_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  labs(caption = "Barcelona neighborhoods: listings") +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot2 <- ggplot(data = Madrid_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  labs(caption = "Madrid neighborhoods: listings") +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

plot3 <- ggplot(data = Valencia_counts) +
  geom_sf(aes(fill = Quantile),
          color = "black",size=.2) +
  labs(caption = "Valencia neighborhoods: listings") +
  ylim(39.405,39.52) +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE","#FFDFA2", "#FFA93F", "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5), "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))
```

@@

The total number of dwellings is available from the Spanish cadastre aggregated by district (districts are groups of neighborhoods). The right column in Figure \ref{fig:all-polygons} shows the percentage of listed dwellings relative to the total number of dwelling by district in each of the three cities. This gives a sense of how active residential real estate markets were in different parts of each city in 2018.

```{r listings-to-properties, include = FALSE}
# Barcelona
Barcelona_prop <- properties_by_district |>
  filter(CITY == "Barcelona") |>
  mutate(proportion = N/N_CADASTRE)

@@

plot4 <- ggplot(data = Barcelona_prop) +
  geom_sf(aes(fill = Quantile),
          color = NA) +
  geom_sf(data = Barcelona_Polygons,
          fill = NA) +
  labs(caption = "Barcelona districts: listings/total dwellings") +
  theme_bw(base_size = 6) +
  scale_fill_manual(values=c("#FFFEDE",
                             "#FFDFA2",
                             "#FFA93F",
                             "#D5610D"))+
  theme(plot.margin=unit(c(0.1, 0.5, 0.1, 0.5),
                         "cm"),
        legend.position = "bottom",
        legend.title = element_blank(),
        legend.key.size = unit(0.2, 'cm'),
        text = element_text(size=4))

@@

# Masking the prices {#anonymizing}

To comply with Spanish regulations, three variables were slightly modified to provide a measure of anonymity. A masking process was applied to asking prices and location (coordinates).

@@

```{r figure-masking-coordinates, echo = FALSE, out.width=c("29%","37%"), out.height="20%",fig.cap="\\label{fig:Anonymizing}(Left) Masking coordinates. Spatial range. (Right) Coordinate displacement in meters (Valencia)",fig.show='hold',fig.align='center'}
knitr::include_graphics(c("EPB_files/points-moved-image.png","EPB_files/coordinates-valencia.png"))
```
